\documentclass{book}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage{makeidx}
\usepackage{index}
\makeindex

\title{Pure Mathematics for Engineers and Scientists}
\author{Anthony Daniell}

\begin{document}

\maketitle 


\chapter*{Preface}


The basic aim is to provide exposure to a solid foundation of mathematics based on proofs.  The exit goals would be:

-Readers will be able to write valid proofs using the standard techniques of modern mathematics.

-Readers will be able to read and understand proofs at some reasonable level.

-Readers will have working knowledge of broad foundations of mathematics, and be able to confidently take dedicated courses in each area, or pursue independent studies.
    
    
From my experience, I certainly never got this stuff explicitly, but on retrospect, it certainly seems like it would help.     


My original concept was that the course would be targeted from someone maybe at the junior or senior level in physics/engineering.  So, the typical prerequisite would be calculus, maybe differential equations, and experience applying this math to physics, engineering problems like in mechanics, electricity and magnetism, etc.  That's essentially my background, so that's why I thought of this cohort.

The main reason is that this person, in my concept, would have some context and exposure to different math, and may be better motivated to wanting to understand the underpinnings.  Also, they may be more interested to branch out into the other topics on the list as these will likely be important as they take more advanced classes in their careers.

However, that being said, I think the main prerequisite is being interested in pure mathematics, perhaps with some future application in mind, but not necessarily.


Topics to be covered:

    Mathematical Foundations - Mathematical logic, basic notation, axiomatic set theory (Zermelo Frankel), Peano Axioms of Arithmetic, Godel theorems.


    Methods of Proof:  Induction, Contradiction, Direct Proofs.  Examples of classic proofs - infinite primes, irrationality of square root of two, etc.  Plenty of practice examples.


    Real Analysis - Study of real numbers, including construction by Dedekind Cuts etc, point set topology, continuity, metric spaces (Basically a compressed version of Rudin or equivalent).


    Modern Algebra:  Groups, Rings, Fields, Transformations, Representation Theory, Modules, Galois Theory


    Non-euclidean geometry and topology:  Various spaces and properties.  Invariants, knot theory, homotopy, homology.


    Number theory:


    Category theory:


    Computational complexity theory:  Finite Automata, lambda calculus, NP and other spaces.
    
\tableofcontents


\include{./chapter1/ch1_MathematicalFoundations}
\include{./chapter2/ch2_MethodsOfProof}

\printindex

\end{document}